---
title: "FRED Real Estate Analysis"
author: "Max Guthrie"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

In this project, data was pooled from FRED (https://fred.stlouis.org/). The dataset was constructed in Excel by using the following variables (Note: the data was reported at a monthy level over 10 years, or 120 months, from January 2012 to January 2022):

* returns on the real estate index = % change in Wilshire US RESI (variable name in Excel: WILLRESIPR_PCH)
* returns on the market index = % change in Wilshire 5000 Full Cap Price Index (variable name in Excel: WILL5000PRFC_PCH)
* risk-free rate = 3-Month Treasury Bill (variable name in Excel: TB3MS)
* Unemployment Rate (variable name in Excel: UNRATE)
* Excess returns on the real estate index = WILLRESIPR_PCH - TB3MS (variable name in Excel: Y)
* excess returns on the market index = WILL5000PRFC_PCH - TB3MS (variable name in Excel: X)
* inflation rate for core CPI = % change in core CPI (variable name in Excel: CPIFESL_PCH)

We import the data into R:

```{r}
FREDdata<-read.table(pipe("pbpaste"), sep="\t", header=TRUE)
```

# Part One
First, we wanted to study the relationship between excess returns on the real estate index (Y) and excess returns on the market index (X) to see if there is any correlation between the two. This can be expressed by the following linear regression model:

$y_t$ = $\alpha$ + $\beta$$x_t$ + $e_t$, where $e_t$ is an error term with mean zero

To find an estimated model for the data, we conducted linear regression in R:

```{r}
reg1=lm(Y~X, data=FREDdata)
summary(reg1)
```

From the above, we get:

$\hat{\alpha}$ = -0.3232
$\hat{\beta}$ = 0.7761

Our estimated model can, therefore, be written as follows:

$\hat{y_t}$ = -0.3232 + 0.7761$x_t$

We can conclude from the data that there is a positive relationship between excess returns on the real estate index and excess returns on the market index. In other words, for every 1% increase in excess returns in the market index, we can expect a 0.7761% increase in excess returns in the real estate index.


# Part Two
Next, we would like to examine whether the excess returns on the real estate index are related to the unemployment rate and core CPI inflation rate. To do so, the following linear regression model was used:

$y_t$ = $\beta_1$ + $\beta_2$$u_t$ + $\beta_3$$\pi_t$ + $w_t$, where $w_t$ is an error term with mean zero

Once again, we run this linear regression model in R:

```{r}
reg2=lm(Y~UNRATE+CPILFESL_PCH, data=FREDdata)
summary(reg2)
```

From the above, we get:

$\hat{\beta_1}$ = -4.1893
$\hat{\beta_2}$ = 0.5186
$\hat{\beta_3}$ = 6.8805

Our estimated model can therefore be written as:

$\hat{y_t}$ = -4.1893 + 0.5186$u_t$ + 6.8805$\pi_t$

This model implies that the unemployment rate and the core CPI inflation rate both have a positive relationship with the excess returns of the real estate index. To explore this further, we would like to see if we can say with 95% certainty that each variable has an effect on the excess returns of the real estate index. To do so, we test the null hypothesis that the parameter of interest equals zero against the alternative hypothesis that the parameter of interest does not equal zero. We use a 5% significance level.

After calculating the t-statistics for each parameter, we were able to reject the null hypothesis that the parameters of interest were equal to zero. In other words, we can say with 95% certainty that the unemployment rate and the inflation rate are directly correlated with the excess returns of the real estate index.
